maxent model avec xvalidation

### 2019-04-03 src/maxent_avec_xvalidation.Rmd
aic injection

```{r pre-install & load packages, include=FALSE}
packages <- c("dismo")
source("../src/ipak.R")
ipak(packages)
```

@@

```{r load pres-background dataset}
presback <- read.csv("./output/bio/presab_nafo_yr.csv", header = TRUE)
colnames(presback)
```

@@

```{r remove unnecessary columns}
presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt))
colnames(presback)
```

@@

```{r}
janprback <- janprback[!is.na(janprback$temp_depth), ]
janprback <- janprback[!is.na(janprback$salinity_depth), ]
janprback <- janprback[!is.na(janprback$o2_depth), ]
janprback <- janprback[!is.na(janprback$chl_surface), ]
janprback <- janprback[!is.na(janprback$nao_sample), ]
janprback <- janprback[!is.na(janprback$nao_winter), ]
```

#create temporally differentiated testing/training data
as per elith and rad.....

```{r} 
#actually you will not use this
noyears <- unique(janprback$year)
train2003 <- subset(janprback, year = "2003")
test2003 <-  subset(janprback, year == "2003")
train2004 <- subset(janprback, year == "2004")
train2005 <- subset(janprback, year == "2005")
train2006 <- subset(janprback, year == "2006")
```

@@

```{r train and test on 2003 inc AIC for reg val}
train2003 <- subset(janprback, year != "2003")
test2003 <-  subset(janprback, year == "2003")
train2003pa <- train2003$occurrence
test2003p <- subset(test2003, occurrence == "1")
test2003b <- subset(test2003, occurrence == "0")
train2003  <- subset(train2003, select = -c(year, occurrence))

mod <- maxent(x = train2003, p = train2003pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE"))
mod
ev <- evaluate(p = test2003p, a= test2003b, model = mod)
ev
threshold(ev)
```

# beta value (regularization) selection

mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "replicatetype=crossvalidate", "replicates=5")) #add to samples because field study with biased data - see https://esajournals.onlinelibrary.wiley.com/doi/10.1890/10-1171.1. note this should be the default but keep here as a note

>>>>

```{r pre-install & load packages, include=FALSE}
packages <- c("dismo", "stringr")
source("../src/ipak.R")
ipak(packages)
```

@@

```{r load pres-background dataset}
presback <- read.csv("../output/bio/presab_nafo_yr.csv", header = TRUE)
colnames(presback)
```

@@

```{r remove unnecessary columns}
presback <- subset(presback, select = -c(cell_id, id, decimalLatitude, decimalLongitude, datecollected, institutioncode, individualcount, depth, resname, originalscientificname, collectioncode, day, nafo_zone, gear, longitude_meters, latitude_meters, total_cell_obs_xy, total_cell_obs_xyt, bottom_depth, XXtotal_cell_obs_xyzt, temp_celsius_depth, temp_celsius_surface, longitude_meters.1, latitude_meters.1, bottom_depth_glorys, longitude_meters.2, latitude_meters.2, cell_id_3d, total_cell_obs_xyzt, cell_id_xyzt))
colnames(presback)
```

@@

```{r jan remove na vals}
janprback <- janprback[!is.na(janprback$temp_depth), ]
janprback <- janprback[!is.na(janprback$salinity_depth), ]
janprback <- janprback[!is.na(janprback$o2_depth), ]
janprback <- janprback[!is.na(janprback$chl_surface), ]
janprback <- janprback[!is.na(janprback$nao_sample), ]
janprback <- janprback[!is.na(janprback$nao_winter), ]
```

@@

```{r train and test on 2003 inc AIC for reg val}
timeslice <- "jan2003"
yr <- str_sub(timeslice, -4)
train <- subset(janprback, year != yr)
trainpa <- train$occurrence
test <-  subset(janprback, year == yr)
testp <- subset(test, occurrence == "1")
testp <- subset(testp, select = -c(year, occurrence))
testb <- subset(test, occurrence == "0")
testb <- subset(testb, select = -c(year, occurrence))
train  <- subset(train, select = -c(year, occurrence))

AIClist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE)

#netam = 0.25
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.25"))

#now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1.... lets try automate some of this..

#count number of non-zero coefficients, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

#betamval <- 0.50
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.50"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 0.75
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=0.75"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 1
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=1"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 1.25
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=1.25"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 1.50
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=1.5"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 1.75
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=1.75"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 2
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=2"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 2.25
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=2.25"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 2.50
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=2.5"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 2.75
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=2.75"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv")
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 3
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=3"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv")
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 3.25
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=3.25"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 3.50
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=3.5"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 3.75
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=3.75"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

betamval <- 4
mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", "betamultiplier=4"))

#count number of non-zero coefficinets, excluding the final four rows
coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE)
coefile <- head(coefile, -4)
coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
nocoeff <- nrow(coefile2)

#no presences used in model (less those used for testing)
maxresults <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"))
notrainingpoints <- maxresults$X.Training.samples
testauc <- maxresults$Test.AUC
trainauc <- maxresults$Training.AUC

#natural log
samplepred <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
samplepred$nat_log <- log(samplepred$Raw.prediction)
sumsamplepred <- sum(samplepred$nat_log)

for (i in 1:nrow(AIClist)){
  if(AIClist$regularization[i] == betamval){
    AIClist$Test.AUC[i] <- testauc
    AIClist$Train.AUC[i] <- trainauc
    AIClist$n[i] <- notrainingpoints
    AIClist$coefficients[i] <- nocoeff
    AIClist$sum.of.logs[i] <- sumsamplepred

  }
}

AIClist$AIC <- ((2*AIClist$coefficients)-(2*AIClist$sum.of.logs))
AIClist$AICc <- (-2*(AIClist$sum.of.logs)+((2*AIClist$coefficients)*(AIClist$n/(AIClist$n - AIClist$coefficients-1))))

write.csv(AIClist, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

```

``

mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE"))
mod

# this is after beta has been selected for
ev <- evaluate(p = testp, a= testb, model = mod)
str(ev)
auc <- ev@auc
cor <- ev@cor
evaluation <- cbind(auc, cor)
write.csv(evaluation, paste0("../output/maxent/models/", timeslice, "/eval.csv"), row.names = FALSE) 
thres <- threshold(ev)
write.csv(thres, paste0("../output/maxent/models/", timeslice, "/threshold.csv"), row.names = FALSE)
```
